### 2023-07-13 paper/sec-dis-pseudodata.tex
sect 2

 The kinematics of a charged-current neutrino DIS event $(x,Q^2, E_\nu)$,
 or alternatively $(x,Q^2, y)$, are uniquely specified by the measurement of three independent
 final-state variables,
 such as $\lp E_\ell,\theta_\ell, W^2\rp$ or $\lp E_\ell,\theta_\ell, E_h \rp$,
 with $E_\ell$ and $\theta_\ell$ being the energy and polar angle of the outgoing
 charged lepton and $E_h$ the total energy of the hadronic final state.
 %
 Most neutrino detectors can only access $E_h$, given that $W^2$ requires
 fully reconstructing this final state.
 %
 A measurement of  $\lp E_\ell,\theta_\ell, E_h \rp$ then fixed the DIS kinematics as:
 \bea
 E_\nu &=& E_h + E_\ell \, , \nonumber \\
 Q^2 &=& 4 ( E_h + E_\ell) E_\ell \sin^2 \lp \theta_\ell/2\rp \, ,\\
 x&=& \frac{4 ( E_h + E_\ell) E_\ell \sin^2 \lp \theta_\ell/2\rp}{2m_N E_h} \, .\nonumber
 \eea
 These relations also indicate how systematic uncertainties affecting the measurement
 of $\lp E_\ell,\theta_\ell, E_h \rp$ translate into uncertainties in the
 reconstructed values of  $(x,Q^2, E_\nu)$, which in turn modify the expected
 event rates in a given bin.

@@

In addition to the statistical uncertainties given by Eq.~(\ref{eq:statistical_uncertainties}),
one needs to also estimate the systematic uncertainties associated to the reconstruction
of the final state leptonic and hadronic variables listed in Table~\ref{tab:FPF_experiments}.
%
Event which would be classified in a given bin in the case
of a perfect detector may end up being classified into a different bin in the presence of systematic
shifts associated to the lepton energy $E_\ell$, lepton scattering angle $\theta_\ell$, and
hadronic energy $E_h$.
%
Hence, for each source of experimental systematic uncertainty we would like to estimate
its impact at the level of event yields and eventually stablish also the associated
correlation matrix.

By means of tree-level DIS kinematics, one can relate the variables describing the neutrino-nucleon
hard-scattering, $(x,Q^2,E_{\nu})$,
with those describing the leptonic and hadronic final state, namely $\lp E_{\ell},\theta_\ell, E_{h} \rp$,
and that correspond to experimentally measured quantities.
%
The mapping between the two sets of variables is 
\bea
E_{\nu} &=& E_\ell + E_h \, ,\nonumber \\
Q^2 &=& 4E_\ell E_{\nu}\sin^2\lp {\theta_\ell/2}\rp  =  4E_\ell \lp E_\ell + E_h  \rp\sin^2\lp {\theta_\ell/2}\rp \, , \\
x &=& \frac{Q^2}{2m_N(E_{\nu} - E_\ell)} = \frac{4E_\ell \lp E_\ell + E_h  \rp\sin^2\lp {\theta_\ell/2}\rp}{2m_NE_{h}} \, , \nonumber
\eea
highlighting how the measurement of $E_{\ell}$, $\theta_\ell$, and $E_{h}$ uniquely specify the hard-scattering
kinematics $(x,Q^2,E_{\nu})$ of the event.
%
Therefore, systematic shifts in the measurement of these final-state variables will modify
the assignment of each event to a given bin of $x$, $Q^2$, and $E_\mu$, introducing a systematic
error in the event yields and hence in the measurement of the double-differential
neutrino scattering cross-section.

In order to determine these bin-by-bin systematic uncertainties,
we extend the calculation delineated in Eq.~(\ref{sec:pseudo-data_generation}) as follows.
%
Let us consider for illustrative purposes the FASER${\nu}$2 case.
%
From Table~\ref{tab:FPF_experiments}, one reads that the corresponding acceptances
and performance parameters for $E_{\ell}$, $\theta_\ell$, and $E_{h}$   are given by
\bea
E_\ell \ge 100~{\rm GeV} \, , \quad \delta E_\ell \sim 30\% \, ,\nonumber \\
E_h \ge 100~{\rm GeV} \, , \quad \delta E_h \sim 30\% \, ,\\
 \theta_\ell \le \tan^{-1}(0.5) \, , \quad \delta\theta_\ell \sim 1~{\rm mrad} \, . \nonumber 
\label{fasernu2systematic_errors}
\eea
In order to translate the systematic uncertainties $\delta E_\ell$, $\delta E_h $,
and $\delta\theta_\ell$ into systematic errors associated to the binned event yields, 
\be
\label{eq:event_yields_systematic_error}
\delta_{\rm sys}^{(E_\ell)} N_{\rm ev}^{(i)} \, ,\quad
\delta_{\rm sys}^{(E_h)} N_{\rm ev}^{(i)}
\, ,\quad
\delta_{\rm sys}^{(\theta_\ell)} N_{\rm ev}^{(i)} \, ,\qquad i=1,\ldots,N_{\rm bin} \, ,
\ee
we generate first a Monte Carlo set of events, denoted by $\mathcal{D}_0$,
composed by $N_{\rm mc} = 10^7$ samples and determine the assignment of each event
to a point in the $\lp x,Q^2,E_{\nu}\rp$ space.
%
By integrating over this sample by means of Eq.~(\ref{MCintegration}) we determine
the binned event yields for this reference $\mathcal{D}_0$ sample.

We then produce a second Monte Carlo sample $\mathcal{D}_1$ starting from the events
of $\mathcal{D}_0$ which are then smeared with Gaussian distributions whose variances
are given by Eq.~(\ref{fasernu2systematic_errors}).
%
The bin assignment of the events in the smeared sample $\mathcal{D}_1$ will
in general be different from those of the baseline sample.
%
The procedure is repeated $M$ times, leading to
$\mathcal{D}_k$ (with $k=1,\ldots,M$) smeared
samples each one leading to a different binned event yields
$ N_{\rm ev}^{(i)(k)}$.
%
Provided the number of samples $M$ is large enough, the standard deviation over samples in this $ N_{\rm ev}^{(i)(k)}$
distribution provides the sought-for estimate of the systematic uncertainties.
%
Furthermore, 
since the systematic errors are treated as uncorrelated among them,
by producing samples where only one source of error is varied at a time
we can determine the values of Eq.~(\ref{eq:event_yields_systematic_error}) in each bin.

Fig.~\ref{fig:percentage_error_elepton}
displays the estimated systematic uncertainties for the  measurements
of the double-differential
neutrino scattering cross-section at FASER$\nu$2.
%
We consider the leading sources of systematic errors,
associated respectively to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$
and to the hadronic energy $E_h$.
%
The size of each source of systematic error is plotted as a function
of the average momentum fraction per bin $\la x\ra$
in three different bins of $Q^2$.
%
We indicate separately the results for neutrino and antineutrino projectiles as well as
those associated to inclusive and to charm production measurements.
%
Systematic uncertainties associated to the hadronic final state energy $E_h$ appear to be the largest.
%
The uncertainties associated to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$ range
between a few percent up to 20\% in the nominal baseline scenario,
and in the case of the latter they are the smallest in the large-$x$ region.
%
Similar overview plots have been produced for the other LHC neutrino experiments
considered in this work.

%-----------------------------------------------------------------------
\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{plots/percentage_errors.pdf}
  \caption{\small Estimated systematic uncertainties for the  measurements
    of the double-differential
    neutrino scattering cross-section at FASER$\nu$2.
    %
    We consider the leading sources of systematic errors,
    associated respectively to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$
    and to the hadronic energy $E_h$.
    %
    The size of each source of systematic error is plotted as a function
    of the average momentum fraction per bin $\la x\ra$
    in three different bins of $Q^2$.
    %
    We indicate separately the results for neutrino and antineutrino projectiles as well as
    those associated to inclusive and to charm production measurements.
  }
  \label{fig:percentage_uncertainties_overview}
\end{figure}
%-----------------------------------------------------------------------

>>>>

 The kinematics of a charged-current neutrino DIS event $(x,Q^2, E_\nu)$,
 or alternatively $(x,Q^2, y)$, are uniquely specified by the measurement of three independent
 final-state variables,
 such as $\lp E_\ell,\theta_\ell, W^2\rp$ or $\lp E_\ell,\theta_\ell, E_h \rp$,
 with $E_\ell$ and $\theta_\ell$ being the energy and polar angle of the outgoing
 charged lepton and $E_h$ the total energy of the hadronic final state.
 %
 Most neutrino detectors can only access $E_h$, given that $W^2$ requires
 fully reconstructing this final state.
 %
 A measurement of  $\lp E_\ell,\theta_\ell, E_h \rp$ then fixed the DIS kinematics as:
 \bea
 E_\nu &=& E_h + E_\ell \, , \nonumber \\
 Q^2 &=& 4 ( E_h + E_\ell) E_\ell \sin^2 \lp \theta_\ell/2\rp \, ,  \label{eq:dis_kinematic_mapping}\\
 x&=& \frac{4 ( E_h + E_\ell) E_\ell \sin^2 \lp \theta_\ell/2\rp}{2m_N E_h} \, .\nonumber
 \eea
 These relations also indicate how systematic uncertainties affecting the measurement
 of $\lp E_\ell,\theta_\ell, E_h \rp$ translate into uncertainties in the
 reconstructed values of  $(x,Q^2, E_\nu)$, which in turn modify the expected
 event rates in a given bin.

@@

In addition to the statistical uncertainties given by Eq.~(\ref{eq:statistical_uncertainties}),
one needs to also estimate the systematic uncertainties associated to the
finite precision in the reconstruction
of the final state leptonic and hadronic variables listed in Table~\ref{tab:FPF_experiments}.
%
For instance, an event which would be classified into a given bin in $(x,Q^2,E_\nu)$ in the case
of a perfect detector may end up being
mis-classified into a different bin in the presence of systematic
shifts associated to the lepton energy $E_\ell$, lepton scattering angle $\theta_\ell$, and
hadronic energy $E_h$, as indicated by  Eq.~(\ref{eq:dis_kinematic_mapping}).

For each possible source of experimental systematic uncertainty, we therefore need to
quantify its impact at the event yield level.
%
In order to determine these bin-by-bin systematic uncertainties,
we extend the calculation delineated in Eq.~(\ref{sec:pseudo-data_generation}) as follows.
%
Let us consider for illustrative purposes the FASER${\nu}$2 case.
%
From Table~\ref{tab:FPF_experiments}, one reads that the detector
acceptance and reconstruction performance is specified by:
\bea
E_\ell \ge 100~{\rm GeV} \, , \quad \theta_\ell \le \tan^{-1}(0.5)\, , \qquad \nonumber\\
\delta E_\ell \sim 30\% \, 
 \quad \delta E_h \sim 30\% \, ,
 \ \, , \quad \delta\theta_\ell \sim 1~{\rm mrad} \, . 
\label{eq:fasernu2systematic_errors}
\eea
To translate the uncertainties in $\delta E_\ell$, $\delta E_h $,
and $\delta\theta_\ell$ into systematic errors associated to the event yields, 
\be
\label{eq:event_yields_systematic_error}
\delta_{\rm sys}^{(E_\ell)} N_{\rm ev}^{(i)} \, ,\quad
\delta_{\rm sys}^{(E_h)} N_{\rm ev}^{(i)}
\, ,\quad
\delta_{\rm sys}^{(\theta_\ell)} N_{\rm ev}^{(i)} \, ,\qquad i=1,\ldots,N_{\rm bin} \, ,
\ee
we generate first a Monte Carlo set of events, denoted by $\mathcal{D}_0$,
composed by $N_{\rm mc} = 10^7$ samples and determine the assignment of each event
to a point in the $\lp x,Q^2,E_{\nu}\rp$ space.
%
By integrating over this sample by means of Eq.~(\ref{MCintegration}) we determine
the binned event yields for this reference $\mathcal{D}_0$ sample.

We then produce a second Monte Carlo sample $\mathcal{D}_1$ starting from the events
of $\mathcal{D}_0$ which are then smeared with Gaussian distributions whose variances
are given by Eq.~(\ref{eq:fasernu2systematic_errors}).
%
The bin assignment of the events in the smeared sample $\mathcal{D}_1$ will
in general be different from those of the baseline sample.
%
The procedure is repeated $M$ times, leading to
$\mathcal{D}_k$ (with $k=1,\ldots,M$) smeared
samples each one leading to a different binned event yields
$ N_{\rm ev}^{(i)(k)}$.
%
Provided the number of samples $M$ is large enough, the standard deviation over samples in this $ N_{\rm ev}^{(i)(k)}$
distribution provides the sought-for estimate of the systematic uncertainties.
%
Furthermore, 
since the systematic errors are treated as uncorrelated among them,
by producing samples where only one source of error is varied at a time
we can determine the systematic errors Eq.~(\ref{eq:event_yields_systematic_error}) in each bin.

Fig.~\ref{fig:percentage_uncertainties_overview}
displays the systematic uncertainties associated
to $E_\ell$, $\theta_\ell$, and $E_h$.
in the  measurements
of the double-differential
neutrino scattering cross-section at FASER$\nu$2.
%
The magnitude of each systematic error is plotted as a function
of the average momentum fraction per bin $\la x\ra$
in three different bins of $Q^2$.
%
We indicate separately the results for neutrino and antineutrino projectiles as well as
those associated to inclusive and to charm production measurements.
%
Systematic uncertainties associated to the hadronic final state energy $E_h$ appear to be the largest.
%
The uncertainties associated to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$ range
between a few percent up to 20\% in the nominal baseline scenario,
and in the case of the latter they are the smallest in the large-$x$ region.
%
The same procedure is adopted for all the experiments listed in Table~\ref{tab:FPF_experiments}.

%-----------------------------------------------------------------------
\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{plots/percentage_errors.pdf}
  \caption{\small Estimated systematic uncertainties for the  measurements
    of the double-differential
    neutrino scattering cross-section at FASER$\nu$2.
    %
    We consider the systematic errors
    associated to the charged lepton energy $E_\ell$ and scattering angle $\theta_\ell$
    and to the hadronic energy $E_h$.
    %
    The size of each source of systematic error is plotted as a function
    of the average momentum fraction per bin $\la x\ra$
    in three different bins of $Q^2$.
    %
    We indicate separately the results for neutrino and antineutrino projectiles as well as
    those associated to inclusive and to charm production measurements.
  }
  \label{fig:percentage_uncertainties_overview}
\end{figure}
%-----------------------------------------------------------------------
